---
title: "Applying Age Adjusted Relaxed/Strict MFPCACox to ADNI data"
author: "Gomon, Daniel"
date: "24/11/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 12)
```

```{r echo = TRUE, eval = TRUE}
rm(list = ls())
load("ADNImFData.Rdata")
```

```{r}
#library(devtools)
#install_github("d-gomon/MFPCA")
library(MFPCA)
```



First we load the data





# Check which variables influence censoring times for IPCW

```{r}
library(survival)
ipcw_dat <- cbind(Y_surv, X_baseline)
ipcw_dat$status <- 1-ipcw_dat$status

coxmod <- coxph(Surv(time, status) ~ AGE + PTGENDER + PTEDUCAT + status.bl + APOE4, data = ipcw_dat)
summary(coxmod)
```
 
Looks like only APOE42 has an influence on censoring times. So we will not use IPCW in validation scores.


# Parameters

Initialize some parameters to perform mFPCA

```{r echo = TRUE, eval = TRUE}
#Choice for PVE and number of basis function in mean smoothing to use.
#We choose for a PVE of 0.93 for every variable and 3 basis functions for mean smoothing
PVE <- 0.93
pve_long_21 <- rep(PVE, length(mdat_21))
uniexp_21 <- vector(mode = "list", length = length(mdat_21))
for(i in 1:length(mdat_21)){
  uniexp_21[[i]] <- list(type = "uFPCA", pve = pve_long_21[i], nbasis_mu = 3, nbasis = 3)
}
#Validation times
taims <- seq(3, 13.5, 0.5)
#Landmark times
lm_times <- seq(3.5, 5, 0.5)
```







# Compare models (Strictly landmarked)

For each landmark time, we compare the following models:

1. Reg_long: NO, Reg_Base: NO, AA: NO
2. Reg_long: YES, Reg_base: NO, AA: NO
3. Reg_long: YES, Reg_base: Yes, AA: NO
4. Reg_long: YES, Reg_base: YES, AA: YES
5. Reg_long: NO, Reg_base: NO, AA: YES



```{r echo = FALSE, eval = FALSE}
StrictLM <- vector(mode = "list", length = length(lm_times))
for(j in 3:length(lm_times)){
  ztemp <- 1
  params = list(mFData = mdat_21, X_baseline = X_baseline, Y_surv = Y_surv,
                landmark_time = lm_times[j], times_pred = taims[which(taims > lm_times[j])],
                age = NULL, FakeLM = FALSE,
                M = 50, uniExpansions = uniexp_21, n_reps = 20, n_folds = 5,
                displaypb = TRUE, n_cores = 4, reg_baseline = FALSE, reg_long = FALSE, type = "scores")
  for(i in 1:5){
    if(i == 2){
      params$reg_long = TRUE
    } 
    if(i == 3){
      params$reg_baseline = TRUE
    }
    if(i == 4){
      params$age = age
    }
    if(i == 5){
      params$reg_long = FALSE
      params$reg_baseline = FALSE
    }
    print(c("landmark time", params$landmark_time))
    print(c("Age Adjusted?", ifelse(is.null(params$age), FALSE, TRUE)))
    print(c("reg_baseline", params$reg_baseline))
    print(c("AgeDM", params$AgeDM))
    StrictLM[[j]][[ztemp]] <- do.call(rcv_mfpccox, params)
    ztemp <- ztemp + 1

  }
}

save(StrictLM, file = "StrictLM.Rdata")
```



```{r}
load("StrictLM.Rdata")
```

Compare all the landmarked methods:

```{r, fig.width = 12}
plot_rcvlist(list(StrictLM[[1]][[1]], StrictLM[[1]][[2]], StrictLM[[1]][[3]], StrictLM[[1]][[4]], StrictLM[[1]][[5]]), legend = c("Long:NO, Base:NO, AA:No", "Long:YES, Base:NO, AA:No", "Long:Yes, Base:Yes, AA:No", "Long:Yes, Base:Yes, AA:Yes", "Long:NO, Base:NO, AA:Yes"), main_title = "Compare Landmarked Methods, LM 3.5 years")

plot_rcvlist(list(StrictLM[[2]][[1]], StrictLM[[2]][[2]], StrictLM[[2]][[3]], StrictLM[[2]][[4]], StrictLM[[2]][[5]]), legend = c("Long:NO, Base:NO, AA:No", "Long:YES, Base:NO, AA:No", "Long:Yes, Base:Yes, AA:No", "Long:Yes, Base:Yes, AA:Yes", "Long:NO, Base:NO, AA:Yes"), main_title = "Compare Landmarked Methods, LM 4 years")

plot_rcvlist(list(StrictLM[[3]][[1]], StrictLM[[3]][[2]], StrictLM[[3]][[3]], StrictLM[[3]][[4]], StrictLM[[3]][[5]]), legend = c("Long:NO, Base:NO, AA:No", "Long:YES, Base:NO, AA:No", "Long:Yes, Base:Yes, AA:No", "Long:Yes, Base:Yes, AA:Yes", "Long:NO, Base:NO, AA:Yes"), main_title = "Compare Landmarked Methods: LM 4.5 years")

plot_rcvlist(list(StrictLM[[4]][[1]], StrictLM[[4]][[2]], StrictLM[[4]][[3]], StrictLM[[4]][[4]], StrictLM[[4]][[5]]), legend = c("Long:NO, Base:NO, AA:No", "Long:YES, Base:NO, AA:No", "Long:Yes, Base:Yes, AA:No", "Long:Yes, Base:Yes, AA:Yes", "Long:NO, Base:NO, AA:Yes"), main_title = "Compare Landmarked Methods: LM 5 years")
```

It seems clear that regularization (LASSO) in this case does better than Landmarked MFPCACox. This is especially visible as we increase Landmark time. The New Age de-meaning process seems to do better than the old one when landmarking.

## Effect of regularization on prediction accuracy

```{r}
plot_rcvlist(list(StrictLM[[1]][[1]], StrictLM[[1]][[2]], StrictLM[[1]][[3]]), legend = c("Long:NO, Base:NO, AA:No", "Long:YES, Base:NO, AA:No", "Long:Yes, Base:Yes, AA:No"), main_title = "Compare Landmarked Methods, LM 3.5 years")

plot_rcvlist(list(StrictLM[[2]][[1]], StrictLM[[2]][[2]], StrictLM[[2]][[3]]), legend = c("Long:NO, Base:NO, AA:No", "Long:YES, Base:NO, AA:No", "Long:Yes, Base:Yes, AA:No"), main_title = "Compare Landmarked Methods, LM 4 years")

plot_rcvlist(list(StrictLM[[3]][[1]], StrictLM[[3]][[2]], StrictLM[[3]][[3]]), legend = c("Long:NO, Base:NO, AA:No", "Long:YES, Base:NO, AA:No", "Long:Yes, Base:Yes, AA:No"), main_title = "Compare Landmarked Methods, LM 4.5 years")

plot_rcvlist(list(StrictLM[[4]][[1]], StrictLM[[4]][[2]], StrictLM[[4]][[3]]), legend = c("Long:NO, Base:NO, AA:No", "Long:YES, Base:NO, AA:No", "Long:Yes, Base:Yes, AA:No"), main_title = "Compare Landmarked Methods, LM 5 years")
```

Regularization seems to improve prediction accuracy when landmarking, especially when more data is available (later landmark times).


## Effect of Age de-meaning with regularized covariates

Compare No AA vs AA when regularizing 

```{r}
plot_rcvlist(list(StrictLM[[1]][[3]], StrictLM[[1]][[4]]), legend = c("No AA", "AA"),
             main_title = "Regularized covariates: compare AgeDM methods (LM 3.5)")

plot_rcvlist(list(StrictLM[[2]][[3]], StrictLM[[2]][[4]]), legend = c("No AA", "AA"),
             main_title = "Regularized covariates: compare AgeDM methods (LM 4)")

plot_rcvlist(list(StrictLM[[3]][[3]], StrictLM[[3]][[4]]), legend = c("No AA", "AA"),
             main_title = "Regularized covariates: compare AgeDM methods (LM 4.5)")

plot_rcvlist(list(StrictLM[[4]][[3]], StrictLM[[4]][[4]]), legend = c("No AA", "AA"),
             main_title = "Regularized covariates: compare AgeDM methods (LM 5)")

```

When regularizing, both AgeDM methods seem to reduce prediction accuracy.


## Effect of Age de-meaning with unregularized covariates

```{r}
plot_rcvlist(list(StrictLM[[1]][[1]], StrictLM[[1]][[5]]), legend = c("Reg: No, AA: No", "Reg: No, AA: Yes"),
             main_title = "Unregularized covariates: compare AgeDM methods (LM 3.5)")

plot_rcvlist(list(StrictLM[[2]][[1]], StrictLM[[2]][[5]]), legend = c("Reg: No, AA: No", "Reg: No, AA: Yes"),
             main_title = "Unregularized covariates: compare AgeDM methods (LM 4)")

plot_rcvlist(list(StrictLM[[3]][[1]], StrictLM[[3]][[5]]), legend = c("Reg: No, AA: No", "Reg: No, AA: Yes"),
             main_title = "Unregularized covariates: compare AgeDM methods (LM 4.5)")

plot_rcvlist(list(StrictLM[[4]][[1]], StrictLM[[4]][[5]]), legend = c("Reg: No, AA: No", "Reg: No, AA: Yes"),
             main_title = "Unregularized covariates: compare AgeDM methods (LM 5)")
```

Again, AgeDM seems to reduce prediction accuracy when landmarking, even without regularization.




# Compare models (Fake landmarking)

Fake Landmarking:

1. Reg_long: NO, Reg_Base: NO, AgeDM: NO (MFPCACox)
2. Reg_long: YES, Reg_base: NO, AgeDM: NO
3. Reg_long: YES, Reg_base: Yes, AgeDM: NO
4. Reg_long: YES, Reg_base: YES, AgeDM: YES
5. Reg_long: NO, Reg_base: NO, AgeDM: YES



```{r echo = FALSE, eval = FALSE}
RelaxLM <- vector(mode = "list", length = length(lm_times))
lm_times <- seq(3.5, 5, 0.5)
for(j in seq_along(seq(3.5, 5, 0.5))){
  ztemp <- 1
  params = list(mFData = mdat_21, X_baseline = X_baseline, Y_surv = Y_surv,
                landmark_time = lm_times[j], FakeLM = TRUE, times_pred = taims[which(taims > lm_times[j])],
                age = NULL, AgeDM = FALSE,
                M = 50, uniExpansions = uniexp_21, n_reps = 20, n_folds = 5,
                displaypb = TRUE, n_cores = 4, reg_baseline = FALSE, reg_long = FALSE, type = "scores")
  for(i in 1:7){
    if(i == 2){
      params$reg_long = TRUE
    } 
    if(i == 3){
      params$reg_baseline = TRUE
    }
    if(i == 4){
      params$age = age
    }
    if(i == 5){
      params$reg_long = FALSE
      params$reg_baseline = FALSE
    }
    print(c("landmark time", params$landmark_time))
    print(c("age", ifelse(is.null(params$age), FALSE, TRUE)))
    print(c("reg_baseline", params$reg_baseline))
    print(c("AgeDM", params$AgeDM))
    RelaxLM[[j]][[ztemp]] <- do.call(rcv_mfpccox, params)
    RelaxLM[[j]][[ztemp]]$params <- params
    ztemp <- ztemp + 1
  
  }
}



save(RelaxLM, file = "RelaxLM.Rdata")
```


```{r}
load("RelaxLM.Rdata")
```








# Figures for article

```{r}
#pdf(file="ADNI_lm35.pdf", width = 16/3, height = 3, pointsize = 5)
setEPS()
postscript(file="ADNI_lm35.eps", width = 16/3, height = 3, pointsize = 5)
plot_rcvlist(list(RelaxLM[[1]][[1]], RelaxLM[[1]][[3]], RelaxLM[[1]][[5]], StrictLM[[1]][[1]], StrictLM[[1]][[3]], StrictLM[[1]][[5]]),
             legend = c("MFPCCox (Relax, No, No)", "LM: Relax, LASSO: Yes, ABC: No", "LM: Relax, LASSO: No, ABC: Yes", "LM: Strict, LASSO: No, ABC: No", "LM: Strict, LASSO: Yes, ABC: No", "LM: Strict, LASSO: No, ABC: Yes"),
             main_title = "MFPCCox vs Alternatives", sub_title = "Landmark time: 3.5 years", lty = c(6, 6, 6, 1, 1, 1))
dev.off()


#pdf(file="ADNI_lm4.pdf", width = 16/3, height = 3, pointsize = 5)
setEPS()
postscript(file="ADNI_lm4.eps", width = 16/3, height = 3, pointsize = 5)
plot_rcvlist(list(RelaxLM[[2]][[1]], RelaxLM[[2]][[3]], RelaxLM[[2]][[5]], StrictLM[[2]][[1]], StrictLM[[2]][[3]], StrictLM[[2]][[5]]),
             legend = c("MFPCCox (Relax, No, No)", "LM: Relax, LASSO: Yes, ABC: No", "LM: Relax, LASSO: No, ABC: Yes", "LM: Strict, LASSO: No, ABC: No", "LM: Strict, LASSO: Yes, ABC: No", "LM: Strict, LASSO: No, ABC: Yes"),
             main_title = "MFPCCox vs Alternatives", sub_title = "Landmark time: 4 years", lty = c(6, 6, 6, 1, 1, 1))
dev.off()


#pdf(file="ADNI_lm45.pdf", width = 16/3, height = 3, pointsize = 5)
setEPS()
postscript(file="ADNI_lm45.eps", width = 16/3, height = 3, pointsize = 5)
plot_rcvlist(list(RelaxLM[[3]][[1]], RelaxLM[[3]][[3]], RelaxLM[[3]][[5]], StrictLM[[3]][[1]], StrictLM[[3]][[3]], StrictLM[[3]][[5]]),
             legend = c("MFPCCox (Relax, No, No)", "LM: Relax, LASSO: Yes, ABC: No", "LM: Relax, LASSO: No, ABC: Yes", "LM: Strict, LASSO: No, ABC: No", "LM: Strict, LASSO: Yes, ABC: No", "LM: Strict, LASSO: No, ABC: Yes"),
             main_title = "MFPCCox vs Alternatives", sub_title = "Landmark time: 4.5 years", lty = c(6, 6, 6, 1, 1, 1))
dev.off()


#pdf(file="ADNI_lm5.pdf", width = 16/3, height = 3, pointsize = 5)
setEPS()
postscript(file="ADNI_lm5.eps", width = 16/3, height = 3, pointsize = 5)
plot_rcvlist(list(RelaxLM[[4]][[1]], RelaxLM[[4]][[3]], RelaxLM[[4]][[5]], StrictLM[[4]][[1]], StrictLM[[4]][[3]], StrictLM[[4]][[5]]),
             legend = c("MFPCCox (Relax, No, No)", "LM: Relax, LASSO: Yes, ABC: No", "LM: Relax, LASSO: No, ABC: Yes", "LM: Strict, LASSO: No, ABC: No", "LM: Strict, LASSO: Yes, ABC: No", "LM: Strict, LASSO: No, ABC: Yes"),
             main_title = "MFPCCox vs Alternatives", sub_title = "Landmark time: 5 years", lty = c(6, 6, 6, 1, 1, 1))
dev.off()
```

```{r}
n <- 4
lmtaims <- c(3.5, 4, 4.5, 5)
 par(oma = c(4,1,1,1), mfrow = c(2, 2), mar = c(2, 2, 1, 1))
 for (i in 1:n){
    plot_rcvlist(list(RelaxLM[[i]][[1]], RelaxLM[[i]][[3]], RelaxLM[[i]][[7]], StrictLM[[i]][[1]], StrictLM[[i]][[3]], StrictLM[[i]][[7]]),
             legend = c("MFPCCox", "LM: Relax, LASSO: Yes, ABC: No", "LM: Relax, LASSO: No, ABC: Yes", "LM: Strict, LASSO: No, ABC: No", "LM: Strict, LASSO: Yes, ABC: No", "LM: Strict, LASSO: No, ABC: Yes"),
             main = paste0("MFPCCox vs Alternatives, Landmark at ", lmtaims[i], " years."))
 }
 par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
 plot(0, 0, type = 'l', bty = 'n', xaxt = 'n', yaxt = 'n')
 legend('bottom',legend = c("Fabricated Metal", "Iron and Steel", "Paper", "Beverages", "Tobacco"), col = c("blue","black", "green", "orange", "pink"), lwd = 5, xpd = TRUE, horiz = TRUE, cex = 1, seg.len=1, bty = 'n')
 # xpd = TRUE makes the legend plot to the figure
```






